---
title: "spaceDeconv "
output: 
  html_document:
      toc: true
      toc_float: true
date: "2022-06-13"
author: Constantin Zackl
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In the following all deconvolution tools for spaceDeconv are tested and run once to showcase and test all tools.


## Imports and Python Setup
```{r imports, echo=TRUE, results="hide", message=FALSE}
library(reticulate)
library(TENxVisiumData) # data
library(TabulaMurisSenisData) # data
library(ggspavis) # visualization
library(SPOTlight)
```
```{r python, echo=TRUE}
# setup python
reticulate::use_condaenv("cell2loc_env")
scanpy <- reticulate::import("scanpy")
reticulate::py_available()
```



## Datasets
We are using two sample datasets from mouse kidneys.


```{r Datasets, echo=TRUE, results='hide', message=FALSE, warning=FALSE, error=FALSE}
# Spatial
sp <- TENxVisiumData::MouseKidneyCoronal()
rownames(sp) <- rowData(sp)$symbol # overwrite EnsemblID

# Single Cell 
sc <- TabulaMurisSenisData::TabulaMurisSenisDroplet(tissues= "Kidney")$Kidney
```

Let's have a look inside the datasets. The console output shows the available cell types and the corresponding number of available cells. The Plot visualizes the number of gene counts per spot and the corresponding Visium Image.


```{r datasetPlots, echo=FALSE, out.width="50%"}
knitr::kable(table(sc$free_annotation, sc$age))

# calculate total counts
colData(sp)$nCounts <- colSums(counts(sp))
x <- "pxl_row_in_fullres" # x coord
y <- "pxl_col_in_fullres" # y coord, from spatialCoords(sp)
ggspavis::plotSpots(sp, x, y, annotate="nCounts", size=2)
ggspavis::plotVisium(sp, x_coord = x, y_coord =  y, spots = FALSE)

```



## Preprocessing

We subset the cells to remove unsuitable cells and NAs. We further subset cells for performance reasons. In a full run this step will be skipped.

This section will further include normalization steps.

```{r preprocessing}
# We are only using data from 18m old cells
sc <- sc[, sc$age=="18m"]
sc <- sc[, !sc$free_annotation %in% c("nan", "CD45")] # remove false data

# downsampling for performance reasons, copied from MarcElosua/SPOTlight
idx <- split(seq(ncol(sc)), sc$free_annotation)
n_cells <- 50
cs_keep <- lapply(idx, function(i) {
    n <- length(i)
    if (n < n_cells)
        n_cells <- n
    sample(i, n_cells)
})
sc <- sc[, unlist(cs_keep)]

```



## SPOTlight

Marker Genes are necessary for SPOTlight. Here i follow the guideline presented on their website. This could be done differently.
A vector of highly variable Genes is also helpful but not required.

```{r SPOTligthParameters}
# SPOTligth parameters
slot <- "cell_ontology_class" # where are the cell types stored, find better term for "slot"
mgs <- NULL # DataFrame of scored Marker genes
# + obviously the single cell and spatial dataset
```



```{r SPOTlight}
colLabels(sc) <- colData(sc)[[slot]] # add Labels

# norm 
sc <- scuttle::logNormCounts(sc) # score marksers

mgs <- scran::scoreMarkers(sc) # can include a subset step, requires logcounts slot/assay?

# this code is from the vignette
# one could perform this differently but the final dataframe needs to be present
mgs_fil <- lapply(names(mgs), function(i) {
  x <- mgs[[i]]
  # Filter and keep relevant marker genes, those with AUC > 0.8
  x <- x[x$mean.AUC > 0.8, ]
  # Sort the genes from highest to lowest weight
  x <- x[order(x$mean.AUC, decreasing = TRUE), ]
  # Add gene and cluster id to the dataframe
  x$gene <- rownames(x)
  x$cluster <- i
  data.frame(x)
})
mgs_df <- do.call(rbind, mgs_fil)

deconvolution <- SPOTlight::SPOTlight(sc, sp, mgs=mgs_df, weight_id = "mean.AUC", verbose=TRUE)
deconvolution <- deconvolution$mat # extract the result
knitr::kable(head(deconvolution[, 4:6])) # just a subset

```
```{r addToObject, echo=TRUE, message = FALSE}
# adding the results to colData
colnames(deconvolution) <- make.names(colnames(deconvolution))

for (i in colnames(deconvolution)){ # each separately
  sp[[i]] <- deconvolution[, i] 
}

# the colData annotation of the spatialExperiment now looks like this:
names(colData(sp))
```
```{r plottingSpotlight, echo=FALSE}
# Visualization of some examples
# x and y from above
ggspavis::plotSpots(sp, x, y, annotate = "kidney.cortex.artery.cell", size=2)
ggspavis::plotSpots(sp, x, y, annotate = "kidney.loop.of.Henle.thick.ascending.limb.epithelial.cell", size=1)
ggspavis::plotSpots(sp, x, y, annotate = "kidney.proximal.convoluted.tubule.epithelial.cell", size=1)
ggspavis::plotSpots(sp, x, y, annotate = "NK.cell", size=2)
```
```{r plottingSpotlightVisium, echo=FALSE}
# You could overlay the image
ggspavis::plotVisium(sp, x_coord = x, y_coord = y, fill="NK.cell", palette = c("red", "white"))
# here white means a higher NK amount, note the red circle in the middle
```



## Cell2location