---
title: "Example"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Example}
  %\VignetteEngine{knitr::knitr}
  %\VignetteEncoding{UTF-8}
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

spacedeconv is a unified interface for the deconvolution of spatial transcriptomics data. In total spacedeconv gives access to `r length(spacedeconv::deconvolution_methods)` deconvolution methods: 
```{r echo=FALSE}
library(spacedeconv)
spacedeconv::deconvolution_methods
```

# Sample Data
For this tutorial we will use one of spacedeconvs sample dataset and the deconvolution algorithm "SCDC".
```{r loadData, message=FALSE}
library(spacedeconv)
library(SpatialExperiment)

data("single_cell_data_3")
data("spatial_data_3")
```


# Preprocessing and Normalization
To get a first insight into the data we type the dataset name into the terminal: 

```{r viewData}
single_cell_data_3

# for performance reasons we are subsampling the data

single_cell_data_3 <- subsetSCE(single_cell_data_3, cell_type_col = "celltype_major", ncells = 180)
```


We can see that the single-cell data is available as a [SingleCellExperiment](https://bioconductor.org/packages/release/bioc/html/SingleCellExperiment.html). In total this dataset includes expression values from 29733 genes and 7986 cells. It can further be seen that cell type information is available in `colData` where cell related annotation is stored. 

As the next step we normalize the data as `cpm` ("Counts Per Million"). The normalization is stored as an additional assay in the single-cell object. 

```{r normalization, warning=FALSE}
single_cell_data_3 <- spacedeconv::preprocess(single_cell_data_3)
spatial_data_3 <- spacedeconv::preprocess(spatial_data_3)

single_cell_data_3 <- spacedeconv::normalize(single_cell_data_3, method = "cpm")
spatial_data_3 <- spacedeconv::normalize(spatial_data_3, method = "cpm")
```
# Build Model

To build a model we need to provide a single cell expression object annotated with cell type information. Let's have a look into the object to see which annotation is available. 
```{r viewData3}
names(colData(single_cell_data_3))
``` 

There are three different resolutions of cell type annotation available. For this tutorial we will choose `"celltype_major"` as cell type  annotation which includes 9 different cell-types. 

Next we can calculate a signature using the annotated single-data and the deconvolution tool "spatialDWLS". 
```{r build_model, message=FALSE, warning=FALSE, eval=FALSE}
signature <- spacedeconv::build_model(
  single_cell_obj = single_cell_data_3,
  cell_type_col = "celltype_major",
  method = "card"
)
```

The signature contains expression values for all celltypes and a subset of distinctly expressed genes. The following example shows the first 10 rows of the signature calculated in the previous step. 
```{r, eval=FALSE}
knitr::kable(round(signature[1:10, ], 4))
```


# Deconvolution
To deconvolute the Visium slide we use the signature we just calculated. Depending on the selected deconvolution tool further parameters need to be passed, in this case `cell_type_col` and `batch_id_col` as well as the scRNA-seq reference dataset. 
```{r deconvolution, message=FALSE, warning=FALSE}
deconv <- spacedeconv::deconvolute(
  spatial_obj = spatial_data_3,
  single_cell_obj = single_cell_data_3,
  cell_type_col = "celltype_major",
  method = "card",
  batch_id_col = "orig.ident",
  assay_sp = "cpm",
  assay_sc = "cpm"
)
```


# Visualization 

The deconvolution results are stored in the spatial object to simplify visualizations. spacedeconv offers 6 visualization functions with different scopes, further explained in the visualization [vignette](spacedeconv_visualization.html).

We can see the available deconvolution results by using `available_results(deconv)` or by accessing the objects `colData`: 

```{r accessColData}
available_results(deconv)
```
For each Spot the estimated cell fraction is available as object annotation and can be used for visualizations. 

```{r visualization, fig.width=12, fig.height=12}
# plot all available results: provide the tool as parameter
spacedeconv::plot_celltype(deconv,
  cell_type = "card",
  density = FALSE,
  smooth = T,
  title_size = 12
)
```

```{r visualization, fig.width=12, fig.height=12}
# ... or plot a specific result
spacedeconv::plot_celltype(deconv,
  cell_type = "card_Cancer.Epithelial",
  density = FALSE,
  smooth = T,
  title_size = 12
)
```
