---
title: "Data Input"
author: "Constantin Zackl"
date: "2022-09-12"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Loading Data}
  %\VignetteEngine{knitr::knitr}
  %\VignetteEncoding{UTF-8}
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Introduction 
SpaceDeconv integrates first- and second-generation deconvolution algorithms for transcriptome data. While first-generation tools deconvolute based on internal precomputed signatures, second-generation deconvolution tools compute cell-type specific expression signatures from annotated single-cell expression data. In the following requirements and workflow details for both types of data are explained in detail.

# Example Data
SpaceDeconv contains 4 different spatial datasets obtained with the 10XVisium platform. Additionally there are 4 matching scRNA-seq datasets available which can be used to calculate custom expression signatures. The data can be loaded like this:

```{r loadingData}
library(SpaceDeconv)
data("single_cell_data_1")
data("single_cell_data_2")
data("single_cell_data_3")
data("single_cell_data_4")
data("spatial_data_1")
data("spatial_data_2")
data("spatial_data_3")
data("spatial_data_4")
```


# Reference Data
To build cell-type specific signatures for second-generation deconvolution SpaceDeconv utilized annotated single-cell reference data. Your single cell data needs to include cell-type annotations and can be one of the following formats: 

* [SingleCellExperiment](https://bioconductor.org/packages/release/bioc/vignettes/SpatialExperiment/inst/doc/SpatialExperiment.html#1_Class_structure) (recommended)
* [AnnData](https://anndata.dynverse.org/)
* [Seurat](https://satijalab.org/seurat/)

Cell type information needs to be available. You can specify the column containing the annotation with the `cell_type_col` parameter. The same applies to batch ID information required for ... with the parameter `batch_id_col`.

```{r annotationParameters, eval=FALSE}
signature <- build_model(single_cell_object,
  method = "dwls",
  cell_type_col = "cell_ontology_class"
)

# ... requires batch_id information as well
sigature <- build_model(single_cell_object,
  method = "",
  cell_type_col = "cell_ontology_class",
  batch_id_col = "batch_id"
)
```


# Spatial Data
Spatially resolved data needs to be available in the SpatialExperiment format. For 10XVisium slides the data can be loaded easily by providing the path to the data folder. More information about data loading and manual object construction can be found in the SpatialExperiment [Vignette](https://bioconductor.org/packages/release/bioc/vignettes/SpatialExperiment/inst/doc/SpatialExperiment.html).

```{r loadSpatialExperiment, eval=FALSE}
spe <- SpatialExperiment::read10xVisium(path_to_directory)
```

Spatial data available as a Seurat Object can be converted into a SpatialExperiment:

```{r seuratConversion, eval=FALSE}
spe <- seurat_to_spatialexperiment(seurat_object)
```

# Normalization
SpaceDeconv offers an additional function for convenient normalization of SpatialExperiments. The normalization is saved in a new assay, so make sure the correct data is used during deconvolution by providing the desired assay with the parameters `assay_sc` and `assay_sp`.

```{r deconvolution, eval = FALSE}
spe <- normalize(spe, method = "cpm")

# make sure to use cpm assay in deconvolution step
deconvolution <- deconvolute(spe, method = "quantiseq", assay_sp = "cpm")
```
