---
title: "AnnData Import"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{spacedeconv anndata tutorial}
  %\VignetteEngine{knitr::knitr}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Setup
This vignette shows how to use `H5AD` `anndata` files as input for the spacedeconv pipeline.
For this, we use the `zellkonverter` package which is also available via `conda`.

Here, we set up `reticulate` before loading the `spacedeconv` and `zellkonverter` libraries to avoid
`reticulate` conflicts between the two packages.

```{r message=FALSE, warning=FALSE}
library(reticulate)
use_condaenv("r-omnideconv", required = TRUE)
invisible(py_config())

library(spacedeconv)
library(zellkonverter)
library(Seurat)
library(SingleCellExperiment)
```


# Load Example Data
First, we load the included example datasets.

```{r data, message=FALSE, warning=FALSE}
data("single_cell_data_3")
data("spatial_data_3")
```


# SingleCellExperiment <-> AnnData Conversion
`zellkonverter` allows us to convert `SingleCellExperiment` objects to `anndata`
and store them as `H5AD` files: [writeH5AD(...)](https://theislab.github.io/zellkonverter/reference/writeH5AD.html)

```{r write_h5ad, message=FALSE, warning=FALSE}
writeH5AD(single_cell_data_3, "single_cell_data_3.h5ad")
```

In case we have the single-cell data only as `H5AD` `anndata` file, we can use `zellkonverter` to 
read it into R as `SingleCellExperiment` object: [readH5AD(...)](https://theislab.github.io/zellkonverter/reference/readH5AD.html)

That way it can be used as input for
the `spacedeconv` pipeline. In this case we read back example data file we just wrote to `H5AD`.

```{r read_h5ad, message=FALSE, warning=FALSE}
single_cell_data_3 <- readH5AD("single_cell_data_3.h5ad", use_hdf5 = FALSE)
```

# SingleCellExperiment <-> Seurat Conversion

`Seurat` allows us to convert `SingleCellExperiment` to `Seurat` objects.

```{r sce_to_seurat, message=FALSE, warning=FALSE}
single_cell_data_3 <- as.Seurat(single_cell_data_3, counts = "counts", data = NULL)
```

If you have single-cell data as `Seurat` object, you can convert it to a `SingleCellExperiment`
to use it as input for `spacedeconv`.

```{r seurat_to_sce, message=FALSE, warning=FALSE}
single_cell_data_3 <- as.SingleCellExperiment(single_cell_data_3, assay = "originalexp")
```

# Spacedeconv Clustering Pipeline
Now we can continue with the spacedeconv pipeline as usual. In this case we will perform
the clustering procedure as an example.

```{r preprocess_normalize, message=FALSE, warning=FALSE}
single_cell_data_3 <- spacedeconv::preprocess(single_cell_data_3)
spatial_data_3 <- spacedeconv::preprocess(spatial_data_3)

spatial_data_3 <- spacedeconv::normalize(spatial_data_3, method = "cpm")
```

```{r build_model, message=FALSE, warning=FALSE}
signature <- spacedeconv::build_model(
  single_cell_obj = single_cell_data_3,
  cell_type_col   = "celltype_major",
  method          = "spatialdwls",
  verbose         = TRUE
)
```

```{r deconvolute, message=FALSE, warning=FALSE}
deconv <- spacedeconv::deconvolute(
  spatial_obj     = spatial_data_3,
  single_cell_obj = single_cell_data_3,
  cell_type_col   = "celltype_major",
  method          = "spatialdwls",
  signature       = signature,
  assay_sp        = "cpm"
)
```

```{r cluster, message=FALSE, warning=FALSE}
clustered <- spacedeconv::cluster(deconv, data = "expression", clusres = 0.5)
```

```{r plot, message=FALSE, warning=FALSE}
spacedeconv::plot_spatial(
  spe     = clustered,
  result  = "cluster_expression_res_0.5",
  title   = "Clustering",
  density = FALSE
)
```
