---
title: "FAQ"
output: rmarkdown::html_vignette
bibliography: references.bib
vignette: >
  %\VignetteIndexEntry{FAQ}
  %\VignetteEngine{knitr::knitr}
  %\VignetteEncoding{UTF-8}
---

# FAQ

### My SpatialExperiment contains ENSEMBL IDs. Can i still perform deconvolution? 
It is possible to perform second-generation deconvolution algorithms when annotated scRNA-seq data containing ENSEMBL IDs is available. All first-gen deconvolution tools require HGNC symbols. You may also check your SpatialExperiment `rowData()` for available annotation to overwrite ENSEMBL IDs. In most cases this can be solved by using the following code: 

```{r, eval = FALSE}
rownames(spe) <- rowData(spe)$symbol
```


### How can i see which deconvolution results are available in my object?
You can use `available_results(object)` or access `names(colData(object))` to see which deconvolution results are available. 

### The spacedeconv plots contain a lot of white spots
When reading raw spatial data the SpatialExperiment contains expression information for all spots, even if they are not located over tissue. Remove those spots with the following code 
```{r, eval=FALSE}
spe <- spe[, colData(spe)$in_tissue == TRUE]
```

### The Visium Slide and the spatial data are not rotated in the same direction
The SpatialExperiment wrapper includes function to rotate/mirror the image. More information can be found in the Visualization Vignette or the SpatialExperiment Documentation. 

### The hexagons in my plots are rotated and don't give an even appearance
Your Visium slide might be aligned incorrectly. Try rotating the image or use `offset_rotation=TRUE` as a parameter in the plotting function.

### Which deconvolution tools are available?
This list includes the methods name and the internally used tokens. More detailed descriptions of each method can be found at the end of this vignette. 
```{r availableMethods}
spacedeconv::deconvolution_methods
```

# References
